---
title: "Host activity before and during infection influences resulting parasite intensities"
author: "David R. Clark"
date: "2024-09-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Set so that long lines in R will be wrapped:
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)


```


```{r packages and loading in dataset, message=FALSE, warning=FALSE}
#Clear the working environment
rm(list=ls())
#Visualization
library(ggplot2)
library(visreg)
library(gridExtra)
source("http://highstat.com/Books/BGS/GAMM/RCodeP2/HighstatLibV6.R")
#(generalized) Linear mixed modeling
library(lme4)
library(glmmTMB)
library(lmodel2)
#Statistcal analysis reporting and model validation
library(performance)
library(car)
library(lmtest)
library(DHARMa)
#Data wrangling
library(dplyr)
library(plyr)
library(tidyverse)
library(tidylog)
library(splancs)

```

\subsection{Visualize some patterns in the raw data}

```{r Loading in completed data set and some quick manipulations before we get into analysis, warning = FALSE}

#Load in dataset from above. This bit of code is meant to save time so people dont have to rerun the entire data parsing and calculating step above. Code will be saved for reproducibility.
IndBehav7 <- read_csv("IndividualBehaviors_20240501_WA.csv")

IndBehav7 <-IndBehav7%>%
rename(Wormaf = Wormbf, AUC = AUC2)
#Subsetting down to Infected individuals
IndBehav7I<-IndBehav7%>%
  filter(Infection==1 & wormJump >= 0)

#getting the mean and SD of worms jumped
summary(IndBehav7I$wormJump)
sqrt(var(IndBehav7I$wormJump))


#Setting some of the factors back to factors
IndBehav7$fishID<-as.factor(IndBehav7$fishID)
IndBehav7$TrialTime<-as.factor(IndBehav7$TrialTime)
IndBehav7$Sex<-as.factor(IndBehav7$Sex)
IndBehav7$Infection<-as.factor(IndBehav7$Infection)
IndBehav7$ContInf<-as.factor(IndBehav7$ContInf)
IndBehav7$Died<-as.factor(IndBehav7$Died)
IndBehav7$Treatment<-as.factor(IndBehav7$Treatment)


#Calculating tissue tolerance for each individual. 
IndBehav7<-IndBehav7%>%
  mutate(ChSMI = LateSMI - PreSMI)%>%
  mutate(TisTol = ChSMI/Totworm)

 #Subsetting down to female and males only to scale their SMI
 #Males
 IndBehavM<-IndBehav7%>%
   filter(Sex=="M")%>%
   mutate(ScPSMI = scale(PreSMI), ScLSMI = scale(LateSMI ))


 #Females
 IndBehavF<-IndBehav7%>%
   filter(Sex=="F")%>%
   mutate(ScPSMI= scale(PreSMI), ScLSMI = scale(LateSMI))

#Combine the two separate dataframes together
 IndBehav8<-rbind(IndBehavF, IndBehavM)

#Scaling some variables to make them biologically comparable and better for model fitting
IndBehav8 <- IndBehav8 %>%
  mutate(
         ScVarvelBef = c(scale(VarvelBef)),
         ScNRatebf = c(scale(NRatebf)),
         ScBehavTol = c(scale(BehavTol)),
         ScTotworm = c(scale(Totworm)),
         ScAUC = c(scale(AUC)),
         ScBehavVig = c(scale(BehavVig)),
         ScChSMI = c(scale(ChSMI)),
         ScTisTol = c(scale(TisTol)))
#Filtering down to only infected and only individuals who have VIE injections for sample size questions
IndBehavind <- IndBehav8 %>%
  distinct(fishID, .keep_all=TRUE)%>%
  filter(Treatment == "VIE")%>%
  filter(Infection=="1")


IndBehavCh<- IndBehav8 %>% 
  filter(TrialTime != "Before") %>%
  mutate(TrialTime = recode(TrialTime, Early = "1",
                            Late = "2",
                            Later = "3"))

IndBehavCh$TrialTime<-as.factor(IndBehavCh$TrialTime)
IndBehavCh$Infection<-as.factor(IndBehavCh$Infection)


IndBehavChI<-IndBehavCh%>%
  filter(Infection == "1")

IndBehavChU<-IndBehavCh%>%
  filter(Infection == "0")


```


<!-- \section{Calculating athe total area of a polygon based on the number of worms and the avergae activity of each individual} -->

<!-- This metric is used as a combination as the tolerance (change in behavior) and resistance (the change in parasites over the course of infection). -->

```{r New Metric of tolerance, warning=FALSE, echo=FALSE, include=FALSE}
#Visually inspecting the ploygons for the change in behavior as parasite load increases
ggplot(IndBehav7, aes(x=Wormaf, y=AvgVel, color=fishID))+
  geom_point()+
  theme_classic()+
  theme(legend.position="none")+
  geom_polygon(aes(fill=fishID, alpha=1/10))

#Creating a dataframe that is missing the NAs for average velocity and Wormaf from fish who died/recovered during infection.
NonNABehav<-IndBehav8 %>%
  drop_na(c(AvgVel, Wormaf))

#Setting Levels as a character
Levels<-as.character(NonNABehav$fishID)

#Subsetting down to only unique variables in fishID/Levels
Levels <- Levels%>%
  unique()



#Calculating the area of those polygons
#Creating empty dataframes and columsn
TotalAreas<-as.data.frame(Levels)
TotalAreas$PolyArea<- NA


#For Loop for calculating the area of the polygon of change in fish behavior with parasite ID
for(i in 1:length(TotalAreas$Levels)){
  current_fish <- TotalAreas$Levels[i]

Convex_hull <- NonNABehav%>%
  filter(fishID == current_fish)%>%
  select(AvgVel, Wormaf)%>%
  slice(chull(Wormaf, AvgVel))

 TotalAreas$PolyArea[i] <- as.matrix(Convex_hull)%>%areapl

 print(current_fish)

}

#Renaming the Levels as fishID for merging
TotalAreas$fishID <- TotalAreas$Levels

#Removing the levels column
TotalAreas <- TotalAreas%>%
  select(-Levels)

#Merging the TotalAreas with our data frame
IndBehav8<-merge(IndBehav8, TotalAreas, by.x=c("fishID"), by.y=c("fishID"), all=TRUE)

#subsetting out before measurements

IndBehav8 <- IndBehav8 %>%
  filter(TrialTime != "Before")%>%
  mutate(ResidPLength = resid(lm(PreLength~Sex)))%>%
  mutate(ResidPLSMI = resid(lm(ScPSMI~ResidPLength)))%>%
  mutate(ScResidPLSMI = c(scale(ResidPLSMI)))%>%
  mutate(ScRPLength = c(scale(ResidPLength)))

#Filtering the data set for individuals who are not controlling for infection.
IndBehavCont<-IndBehav8%>%
  filter(ContInf == "0" | ContInf == "1")

#Filtering to only infected individuals for tolerance metrics
IndBehavI<-IndBehav8%>%
  filter(Infection == "1")%>%
  distinct(fishID, .keep_all=TRUE)


```

\subsection{What hypotheses we want to test with these data and what data we can use to test them?}

- Does host behavioral vigor trade-off with their ability to resist parasite infection?

- Does higher host activity during infection lead to higher parasite infection?

```{r more specific visualization of explanatory variables mentioned above, warning = FALSE, include=FALSE}

pairs(~AvgVel+Sex+Infection+AUC+NRatebf+VarVel+BehavTol+ScPSMI+BehavVig+TisTol+ResidPLength+Treatment, lower.panel=panel.smooth, diag.panel=panel.hist, 
upper.panel=panel.cor, data=IndBehav8)

```




\subsection{What factors are important for host infection intensity over the course of infection and do is there any sexual variation in host infection intensity?}

\subsection{Visually inspection of the explanatory variables that will be used in the analyses}
```{r vsiaully inspecting covariates for the tolerance and vigor model, warning=FALSE}

pairs(~AUC+Sex+ScPSMI+ScVarvelBef+ScBehavTol+ScBehavVig+PreLength+Treatment, lower.panel=panel.smooth, diag.panel=panel.hist, 
upper.panel=panel.cor, data=IndBehav8)

```

\subsubsection{Description, development, and fitting of linear model for the analysis}

We will use a linear model to analyze how infection intensity differs by sexual variation and other important host traits. Given each host only has one infection intensity measure, we do not need the fishID random effect used in previous models.

* Deterministic
 + $AUC_{det}$ = a + $b_{1}$Sex + $b_{2}$ScBehavVig + $b_{3}$ScVarVelBef + $b_{4}$ScResidPLSMI + $b_{5}$ScRPLength + $b_{6}$Sex:ScBehavVig + $b_{7}$Sex:ScResidPLSMI

* Stochastic
  + AUC ~ *N*($AUC_{det}$, $\sigma^{2}$) 

  
* Fixed
  + Sex 
  + Scaled behavioral vigor 
  + Scaled Pre-infection SMI
  + Scaled variance in velocity before infection 
  + Scaled Residuals of Length and Sex 
  + VIE treatment
  + Interaction between Sex and behavioral vigor
  + Interaction between sex and body condition
  
  
```{r linear model for Infection Integral}
#Scaling our pre-infection sclaed mass index
IndBehavI$ScPSMI<-as.numeric(IndBehavI$ScPSMI)
#Fit a linear model for infection integral
#Fitting a glm because we only have one measure of beahvioral vigor and infection integral 
AUCLM<-glm(AUC~Sex+ScBehavVig+ScVarvelBef+ScRPLength+Treatment+wormJump+Sex:ScRPLength, family=Gamma(link="log"), IndBehavI)

#Summary to see the relationship of the variables. 
summary(AUCLM)

```

\subsubsection{Validate that the model fits well and there are no problems}

```{r model validation for the AUC model, warning=FALSE}

#Using the check_model function from the perforamnce package to check the model validation
check_model(AUCLM)


#Using the Dharma package to check quantile residuals
#First simulating the quantile residuals
sim_residuals_AUCLM <-simulateResiduals(AUCLM, 1000)
#Plotting the quantile residuals to test how quantile residuals look
plot(sim_residuals_AUCLM) 
#Testing for dispersion
testDispersion(sim_residuals_AUCLM)

```


\subsubsection{Testing the significance of factors in our model}

```{r checking significance of fixed effects for AUC model, warning=FALSE}

Anova(AUCLM, type=3)

```

\subsubsection{Visualize the important explanatory factors for infection intensity}

```{r visualing the pattern between Infection integral and behav Vig, warning=FALSE}
#Behavioral Vigor graph
InfIntbyVig=visreg(AUCLM, scale="response", "ScBehavVig", partial=T, gg=TRUE)+
  theme_classic()+
  theme(legend.position="none")+
  ylab("Infection intensity")+
  xlab("Scaled Behavioral Vigor")+
  theme(text=element_text(size=18))
#Print the graph
print(InfIntbyVig)
```

```{r visualing the pattern between Infection integral and Variance in velocity, Warning=FALSE}
#Behavioral Vigor graph
#Extracting fit and residuals from model
InfIntbyLen=visreg(AUCLM, scale="response", "ScRPLength", "Sex", partial=T, plot=FALSE, overlay=T)
#Extracting fit
InfIntbyLenfit<-InfIntbyLen$fit
#Extracting residuals
InfIntbyLenres<-InfIntbyLen$res
cpsex=c("darkseagreen4","lightsalmon3")
InfIntbyLengraph<-ggplot()+
  geom_smooth(data=InfIntbyLenfit, aes(x=ScRPLength, y=visregFit, group=Sex, color=Sex))+
  geom_ribbon(data=InfIntbyLenfit, aes(x= ScRPLength, y= visregFit, ymin = visregLwr, ymax = visregUpr,fill=Sex), alpha=0.5)+
  geom_point(data=InfIntbyLenres, aes(x=ScRPLength, y=visregRes, group=Sex, color=Sex))+
  theme_classic()+
  theme(legend.position="none")+
  ylab("Infection intensity")+
  xlab("Scaled Residual body length")+
  theme(text=element_text(size=18))+
  scale_color_manual(values=cpsex)+
  scale_fill_manual(values=cpsex)+
  ylim(0,7000)
#Print the graph
print(InfIntbyLengraph)
```

```{r visualing the pattern between Infection integral and SMI by sex, Warning=FALSE}
#Sex by SMI graph
#Extracting the fit and residuals
InfIntbySex=visreg(AUCLM, scale="response", "Sex", partial=T, plot=FALSE)
#Extract fit
InfIntbySexFit <- InfIntbySex$fit
#Extract residuals
InfIntbySexRes <- InfIntbySex$res
#Add colorscheme by sex
cpsex=c("darkseagreen4","lightsalmon3")
#Plot the graph
InfSexgraph<-ggplot(data=InfIntbySexRes, aes(x=Sex, y=visregRes))+
  geom_boxplot(aes(fill=Sex))+
  geom_jitter(alpha=0.5)+
  theme_classic()+
  theme(legend.position="none")+
  ylab("Infection intensity")+
  xlab(" ")+
  theme(text=element_text(size=18))+
  scale_fill_manual(values=cpsex)+
  scale_x_discrete(breaks=c("F","M"),
      labels=c("Females", "Males"))
#Loading in gridExtra for multiple graphs in one image
library(gridExtra)

#
InfIntgraphs<-grid.arrange(InfIntbyLengraph, InfIntbyVig, nrow = 1)

```

\subsubsection{Fitting posthoc model to test for pattern within sex}

```{r linear model for Infection integral}
#Subsetting down to females only
IndBehavIF <- IndBehavI %>%
  filter(Sex == "F")

#Subsetting down to males only
IndBehavIM <- IndBehavI %>%
  filter(Sex == "M")

#Fit a linear model for infection intensity by sex
#Female model
AUCLMF<-glm(AUC~ScBehavVig+ScVarvelBef+ScResidPLSMI+ScRPLength+Treatment, family=Gamma(link="log"), IndBehavIF)
#Male model
AUCLMM<-glm(AUC~ScBehavVig+ScVarvelBef+ScResidPLSMI+ScRPLength+Treatment, family=Gamma(link="log"), IndBehavIM)

#Summary to see the relationship of the variables. 
#Females
summary(AUCLMF)
#Males
summary(AUCLMM)
#Testing for significance for females
Anova(AUCLMF, type=3, test="LR")
#Testing for significance for males
Anova(AUCLMM, type=3, test="LR")

```

\subsubsection{Examining change in activity for only infected individuals}


```{r subsetting the data, setting some factors, and quickly visualizing some of our vairables}
#Setting some variables as factors to confirm they are this way for the analysis
IndBehavChI$Sex<-as.factor(IndBehavChI$Sex)
IndBehavChI$Infection<-as.factor(IndBehavChI$Infection)
IndBehavChI$Treatment<-as.factor(IndBehavChI$Treatment)

#Subsetting our dataframe to relevant data and making some 
IndBehavChI<-IndBehavChI %>%
  mutate(ScWormaf = scale(Wormaf)) %>%
  mutate(ScChBehav = scale(ChBehav)) 


pairs(~ChBehav+Sex+Infection+TrialTime+PreSMI+AUC+Treatment+NRatebf+Wormaf, lower.panel=panel.smooth, diag.panel=panel.hist, 
upper.panel=panel.cor, data=IndBehavChI)
cpsex=c("darkseagreen4","lightsalmon3")
```



We will use a linear mixed model to analyze how Change in activity differs by sexual variation and how infected an individual fish was. FishID is included as a random term to allow for non-independence of individuals due to multiple measurements per individual across time.

* Deterministic
 + $Log_{10}$($InfInt_{det} + 1$) = a + $b_{1}$TrialTime + $b_{2}$ChBehav + $b_{3}$Sex + $b_{4}$ScPSMI + $b_{5}$Treatment + $b_{6}$wormJump + $b_{7}$ScChSMI + $b_{8}$BehavVig + $a_{i}$

* Stochastic
  + ChVel ~ *N*($Log_{10}$($InfInt_{det} + 1$), $\sigma^{2}$) 
  + $a_{i}$ ~ *N*(0, $\sigma^{2}_{fishID}$) 
  
* Fixed
  + TrialTime
  + Change in behavior between time points
  +  Sex 
  + Scaled pre-infection scaled mass index
  + VIE treatment of the fish
  + number of worms that initially started the infection
  + Scaled change in scaled mass index
  + Behavior vigor of each fish
  
* Random
  + fishID


```{r fitting the infection intensity model, warning=FALSE}

#Fit a linear model for checking what explanatory factors are important for Variance in activity
#Note this is a linear mixed model because we have multiple measures per fish and therefore, need to account for non-independence between measures. 
InfIntLMInf<-lmer(log10(Wormaf+1)~
                   Sex+
                   ScPSMI+
                   ScChSMI+
                   TrialTime+
                   ScBehavVig+
                   Treatment+
                   ChBehav+
                   wormJump+
                   (1|fishID),
                 IndBehavChI)


#Summary to see the relationship of the variables. 
summary(InfIntLMInf)
```

\subsubsection{Validate that the model fits well and there are no problems}


```{r model validation for the Change in activity model split for infection, warning=FALSE}

#Using the check_model function from the perforamnce package to check the model validation

check_model(InfIntLMInf,check=c("qq","normality", "homogeneity"))

#Using the Dharma package to check quantile residuals
#First simulating the quantile residuals
sim_residuals_InfIntLMInf <-simulateResiduals(InfIntLMInf, 1000)
#Plotting the quantile residuals to test how quantile residuals look
plot(sim_residuals_InfIntLMInf) 
#Testing for dispersion
testDispersion(sim_residuals_InfIntLMInf)

#There are some problems with this model validation. It doesnt look model breaking but definitely should look at other model error structures to resolve the issues. 

```

 \subsubsection{Testing the significance of factors in our model using a Kenward-Rodgers F test for infected only infection intensity model}


```{r using anova to perform a kenward-Rodgers F test for infection intensity, warning=FALSE}

#F test to test for signficance of slope of variables
Anova(InfIntLMInf, test="F", type=3)


```

\subsubsection{Visualize the important explanatory factors for infection intensity for infected individuals}


```{r visualizing Change in activity by infeciton intensity, warning=FALSE}

#Change in behavior by 
#Extracting fit and residual
ChVelbyRate=visreg(InfIntLMInf, scale="response", "ChBehav", partial=T, overlay=TRUE, plot=FALSE)
#Extracting fit
ChVelbyRatefit<-ChVelbyRate$fit
#Extracting residual
ChVelbyRateres<-ChVelbyRate$res
#Constructing the graph 
ChVelbyRategraph<-ggplot(ChVelbyRateres, aes(x=ChBehav, y=visregRes))+
  geom_smooth(method="lm")+
  geom_point()+
  theme_classic()+
  theme(legend.position="none")+
  ylab("Log (number of parasites + 1)")+
  xlab("Change in activity")+
  theme(text=element_text(size=18))

#Print the graph
print(ChVelbyRategraph)


```


```{r fitting post hoc model for Infected males and females}

#Subset to only infected females
IndBehavChIF <- IndBehavChI %>%
  filter(Sex == "F" & Infection == "1")%>%
  mutate(ScWormaf = scale(Wormaf))

#Subset to only infected males
IndBehavChIM <- IndBehavChI %>%
  filter(Sex == "M" & Infection == "1")%>%
  mutate(ScWormaf = scale(Wormaf))


#Fit linear mixed models to post hoc check for significance for male and female slopes
ChVelLMInfF<-lmer(ChBehav~ScPSMI+ScChSMI+TrialTime+ScBehavVig+Treatment+ScNRatebf+(1|fishID), IndBehavChIF)

#Fit linear mixed models to post hoc check for significance for male and female slopes
ChVelLMInfM<-lmer(ChBehav~ScPSMI+ScChSMI+TrialTime+ScBehavVig+Treatment+ScNRatebf+(1|fishID), IndBehavChIM)

#Summary to see the relationship of the variables for female model. 
summary(ChVelLMInfF)

#Summary to see the relationship of the variables for male model. 
summary(ChVelLMInfM)

#Anova to test for significance  for females
Anova(ChVelLMInfF, test="F", type=2)

#Anova to test for significance  for males
Anova(ChVelLMInfM, test="F", type=2)


```
