---
title: "TolVigandResil_20230223"
author: "David R. Clark, Jason Walsman, Rachael Kramp, and Jessica Stephenson"
date: "2024-02-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Set so that long lines in R will be wrapped:
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)
```



```{r packages and loading in dataset, message=FALSE, warning=FALSE}
#Clear the working environment
rm(list=ls())
#Visualization
library(ggplot2)
library(visreg)
source("http://highstat.com/Books/BGS/GAMM/RCodeP2/HighstatLibV6.R")
#(generalized) Linear mixed modeling
library(lme4)
library(glmmTMB)
library(lmodel2)
#Statistcal analysis reporting and model validation
library(performance)
library(car)
library(lmtest)
library(DHARMa)
#Data wrangling
library(dplyr)
library(plyr)
library(tidyverse)
library(tidylog)

```

```{r load in data, warning=FALSE}

#Loading in data set
IndBehav<- read_csv("VIEBehavior_20210913_V2.csv", 
    col_types = cols(fishID = col_character(), 
         Infection = col_character(), InfDate = col_character(), 
         BehavGroup = col_character(), dayofinf = col_character(), 
         CountDay = col_character(), Behavdate = col_character(), 
         TOD = col_character(), Duration=col_double(),Velocity=col_double(),Distance=col_double()))

#Filtering Data set so we can perform our analysis
IndBehav1<-IndBehav %>%
  #Removing columns we dont need
  select(-c(Blind.ID, fishID...19, TrialTime...31, AUC2...37, InfLength, Recov...40,Ratebf...48, LateLength, dayofinf, Countintial, Notes, InfWeight))%>%
  #Renaming some of the columns with funky names for ease of coding and naming
  rename(fishID = fishID...2, AUC=AUC2...21, TrialTime=TrialTime...23, Ratebf=Ratebf...24, Recov=Recov...25)

```
```{r calculating the body metrics, warning=FALSE}


IndBehav1<-IndBehav %>%
  select(-c(Blind.ID, fishID...19, TrialTime...31, AUC2...37, InfLength, Recov...40,Ratebf...48, LateLength, dayofinf, Countintial, Notes, InfWeight))%>%
  rename(fishID = fishID...2, AUC=AUC2...21, TrialTime=TrialTime...23, Ratebf=Ratebf...24, Recov=Recov...25)

#####This code is just to examine how many fish cleared their infection (13)####
# IndBehavRec<- IndBehav1 %>%
#   filter(Recov==1)%>%
#   filter(Fish==1)


#####This code is just to examine how many fish died during their infection (2)#####
 # IndBehavDead<- IndBehav1 %>%
 #   filter(Died==1)%>%
 #   filter(Fish==1)

FemaleOnly<-IndBehav1 %>%
  filter(Sex == 'F')
  
MaleOnly<-IndBehav1 %>%
  filter(Sex == 'M')


#Calculating SMI metrics
#####Preinfection SMI#####

#PreinfectionSMI for females
#To calculate the SMI we take the OLS slope 
lmodel2(log(as.numeric(PreWeight)+1)~log(as.numeric(PreLength)+1), data=FemaleOnly)
#Take the median length from summary
summary(FemaleOnly$PreLength)
#Adding PreSMI to Dataframe
FemaleOnly <- FemaleOnly %>%
  group_by(fishID)%>%
  mutate(PreSMI = PreWeight*((20.20/PreLength)^0.4206982))

#PreinfectionSMI for males
lmodel2(log(as.numeric(PreWeight)+1)~log(as.numeric(PreLength)+1), data=MaleOnly)
#Take the median length from summary
summary(MaleOnly$PreLength)
#Adding PreSMI to Dataframe
MaleOnly <- MaleOnly %>%
  group_by(fishID)%>%
  mutate(PreSMI = PreWeight*((15.30/PreLength)^0.1938185))


#####Late Infection SMI####

#Late infection SMI for females
#Take OLS slope from this model
lmodel2(log(as.numeric(LateWeight)+1)~log(as.numeric(PreLength)+1), data=FemaleOnly)
#Take the median length from summary
summary(FemaleOnly$PreLength)
#Adding PreSMI to Dataframe
FemaleOnly <- FemaleOnly %>%
  group_by(fishID)%>%
  mutate(LateSMI = PreWeight*((20.10/PreLength)^-0.5976124))

#Late Infection SMI for males
#Take OLS slope from this model
lmodel2(log(as.numeric(LateWeight)+1)~log(as.numeric(PreLength)+1), data=MaleOnly)
#Take the median length from summary
summary(MaleOnly$PreLength)
#Adding PreSMI to Dataframe
MaleOnly <- MaleOnly %>%
  group_by(fishID)%>%
  mutate(LateSMI = PreWeight*((15.30/PreLength)^-0.472241))
#####

#Bringing the male and female dataframes together to create our overall dataframe
IndBehav2<- rbind(FemaleOnly, MaleOnly)

#View(IndBehav2)
```

We are calculating behavioral tolerance through two metrics. First is a measure of behavioral tolerance using linear mixed models using slope random effects for fishID. This allows us to extract each individuals change in activity with parasite burden and therefore allow us to calculate the tolerance (fishID slope) for each individual. This also allows us to use the random effect intercept for fishID as a measure of beahvioral vigor. The second way we could caluclate these metrics are a measure of change in behavior between two points of behavior (often referred to as point tolerance, CITATION FOR THIS).  This is calculated in teh second half of the code as the change in behavior between pre-infection and early, late, and later points of infection. 

```{r calculating the tolerance metrics, warning=FALSE}

#Calulating the behavioral tolerance metrics were interested in for our analysis

#Random slope from random effects model
#Random effects model with random slope for each fish by worm burden prior to measuring its activity. This slope is the tolerance an individual has to changing its behavior as infection increases. The slope of the behavior is the individuals behavioral vigor, or its pre-infection behavior
BehavTolLM<-lmer(Velocity~(Wormbf|fishID), IndBehav2)
summary(BehavTolLM)
#Extracting the intercept and slope from the linear mixed model
betaTol <- data.frame(coef(BehavTolLM)$fishID)
#Renaming the slope and intercept to align with our metrics
betaTol <- betaTol %>%
  rename(BehavTol = Wormbf, BehavVig = X.Intercept.)
View(betaTol)
#Creating a fishID column for combination with future dataframe
betaTol<-cbind(fishID=rownames(betaTol), betaTol)
#Narrowing our overall data set to one row per individual
IndBehav3 <- IndBehav2 %>%
  group_by(fishID)%>%
  filter(TOD == 2) %>%
  select(-c(InfDiff, Fish, Ratebf, Behavdate, Duration, Distance, comp3, Totworm)) %>%
  pivot_wider(names_from = TrialTime, values_from = c(Velocity, Wormbf)) 

#subset down to infected individuals only and calculate point tolerance metrics for each timeframe. 
IndBehav3Inf <- IndBehav3 %>%
  group_by(fishID)%>%
  filter(Infection == 1) %>%
  mutate(ETol = Velocity_Early - Velocity_Before,
         LTol = Velocity_Late - Velocity_Before,
         LtrTol = Velocity_Later - Velocity_Before
         )%>%
  pivot_longer(cols=starts_with("Wormbf"),
               names_to = "TrialTime",
               names_prefix = "Wormbf_",
               values_to = "Wormbf")
# %>%
#   pivot_longer(cols = starts_with("Velocity"),
#     names_to = "TrialTime",
#     names_prefix = "Velocity_",
#     values_to = "Velocity"
#   )

#Merging the tolerance and vigor metrics 
IndBehav3Inf1<-merge(IndBehav3Inf, betaTol, by.x="fishID", by.y = "fishID")



#Subset down to uninfected individuals and calculate their point tolerance metrics
IndBehav3Unf <- IndBehav3 %>%
  filter(Infection == 0)%>%
  mutate(ETol = Velocity_Early - Velocity_Before,
         LTol = Velocity_Late - Velocity_Before,
         LtrTol = Velocity_Later - Velocity_Before
         )%>%
  group_by(fishID)%>%
  pivot_longer(cols = starts_with("Wormbf"),
      names_to = "TrialTime2",
     names_prefix = "Wormbf_",
     values_to = "Wormbf"
   )%>%
  pivot_longer(cols = starts_with("Velocity"),
  names_to = "TrialTime",
  names_prefix = "Velocity_",
  values_to = "Velocity"
  )
#Creating the BehavTol and BehavVig columns for the uninfected and calculating the metric for uninfected individual so we can combine these dataframes together
IndBehav3Unf <- IndBehav3Unf%>%
  group_by(fishID)%>%
  add_column(BehavTol=0)%>%
  mutate(BehavVig = mean(Velocity[IndBehav3Unf$TrialTime == "Before"]))

#Rbinding our infected and uninfected dataframes together into one overall dataframe
IndBehav4<-rbind(IndBehav3Unf,IndBehav3Inf1)

IndBehav4 %>% 
  arrange(fishID, TrialTime2) %>%
  group_by(fishID, TrialTime2) %>%
  mutate(id=1:n()) -> response2

IndBehav5<-response2 %>%
  filter(id == "1")%>%
  select(-c(Velocity, TrialTime,id))
```



```{r calculating whether individuals are recovering during infeciton}
#Quick veiwing of data frame so we can manipulate it in the way we want
View(IndBehav5)

#Making the growth rate columns into one column

IndBehav6 <- IndBehav5 %>%
  group_by(fishID)%>%
  select(-c(rate012, rate1216))%>%
  rename(ratebf_Early = rate05, ratebf_Late = rate512, ratebf_Later = rate18end)%>%
  add_column(ratebf_Before = 0)%>%
  pivot_longer(cols = starts_with("ratebf"),
      names_to = "TrialTime3",
     names_prefix = "ratebf_",
     values_to = "Ratebf"
   )
  


View(IndBehav6)
```

