---
title: "Behavioral Tolerance, Vigor, and resilience"
author: "David R. Clark, Jason Walsman, Rachael Kramp, and Jessica Stephenson"
date: "2024-03-25"
output:
  pdf_document:
    toc: yes
    number_sections: yes
  editor_options:
    chunk_output_type: console
  word_document:
    toc: yes
header-includes:
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Set so that long lines in R will be wrapped:
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

\section{Introduction}

\subsection{Overall summary}

\subsection{Quick experimental summary}

\subsection{Load in packages needed for analysis}

```{r packages and loading in dataset, message=FALSE, warning=FALSE}
#Clear the working environment
rm(list=ls())
#Visualization
library(ggplot2)
library(visreg)
source("http://highstat.com/Books/BGS/GAMM/RCodeP2/HighstatLibV6.R")
#(generalized) Linear mixed modeling
library(lme4)
library(glmmTMB)
library(lmodel2)
#Statistcal analysis reporting and model validation
library(performance)
library(car)
library(lmtest)
library(DHARMa)
#Data wrangling
library(dplyr)
library(plyr)
library(tidyverse)
library(tidylog)

```

\subsection{Load in dataframe}

```{r load in data, warning=FALSE}

#Loading in data set
IndBehav<- read_csv("VIEBehavior_20210913_V2.csv", 
    col_types = cols(fishID = col_character(), 
         Infection = col_character(), InfDate = col_character(), 
         BehavGroup = col_character(), dayofinf = col_character(), 
         CountDay = col_character(), Behavdate = col_character(), 
         TOD = col_character(), Duration=col_double(),Velocity=col_double(),Distance=col_double()))

#Filtering Data set so we can perform our analysis
IndBehav1<-IndBehav %>%
  #Removing columns we dont need
  select(-c(Blind.ID, fishID...19, TrialTime...31, AUC2...37, InfLength, Recov...40,Ratebf...48, LateLength, dayofinf, Countintial, Notes, InfWeight))%>%
  #Renaming some of the columns with funky names for ease of coding and naming
  rename(fishID = fishID...2, AUC=AUC2...21, TrialTime=TrialTime...23, Ratebf=Ratebf...24, Recov=Recov...25)

```
\subsection{Calculate metrics needed for further analyses}

\subsubsection{Calculating all body condition metrics}

```{r calculating the body metrics, warning=FALSE}


IndBehav1<-IndBehav %>%
  select(-c(Blind.ID, fishID...19, TrialTime...31, AUC2...37, InfLength, Recov...40,Ratebf...48, LateLength, dayofinf, Countintial, Notes, InfWeight))%>%
  rename(fishID = fishID...2, AUC=AUC2...21, TrialTime=TrialTime...23, Ratebf=Ratebf...24, Recov=Recov...25)

#####This code is just to examine how many fish cleared their infection (13)####
# IndBehavRec<- IndBehav1 %>%
#   filter(Recov==1)%>%
#   filter(Fish==1)


#####This code is just to examine how many fish died during their infection (2)#####
 # IndBehavDead<- IndBehav1 %>%
 #   filter(Died==1)%>%
 #   filter(Fish==1)

FemaleOnly<-IndBehav1 %>%
  filter(Sex == 'F')
  
MaleOnly<-IndBehav1 %>%
  filter(Sex == 'M')


#Calculating SMI metrics
#####Preinfection SMI#####

#PreinfectionSMI for females
#To calculate the SMI we take the OLS slope 
lmodel2(log(as.numeric(PreWeight)+1)~log(as.numeric(PreLength)+1), data=FemaleOnly)
#Take the median length from summary
summary(FemaleOnly$PreLength)
#Adding PreSMI to Dataframe
FemaleOnly <- FemaleOnly %>%
  group_by(fishID)%>%
  mutate(PreSMI = PreWeight*((20.20/PreLength)^0.4206982))

#PreinfectionSMI for males
lmodel2(log(as.numeric(PreWeight)+1)~log(as.numeric(PreLength)+1), data=MaleOnly)
#Take the median length from summary
summary(MaleOnly$PreLength)
#Adding PreSMI to Dataframe
MaleOnly <- MaleOnly %>%
  group_by(fishID)%>%
  mutate(PreSMI = PreWeight*((15.30/PreLength)^0.1938185))


#####Late Infection SMI####

#Late infection SMI for females
#Take OLS slope from this model
lmodel2(log(as.numeric(LateWeight)+1)~log(as.numeric(PreLength)+1), data=FemaleOnly)
#Take the median length from summary
summary(FemaleOnly$PreLength)
#Adding PreSMI to Dataframe
FemaleOnly <- FemaleOnly %>%
  group_by(fishID)%>%
  mutate(LateSMI = PreWeight*((20.10/PreLength)^-0.5976124))

#Late Infection SMI for males
#Take OLS slope from this model
lmodel2(log(as.numeric(LateWeight)+1)~log(as.numeric(PreLength)+1), data=MaleOnly)
#Take the median length from summary
summary(MaleOnly$PreLength)
#Adding PreSMI to Dataframe
MaleOnly <- MaleOnly %>%
  group_by(fishID)%>%
  mutate(LateSMI = PreWeight*((15.30/PreLength)^-0.472241))
#####

#Bringing the male and female dataframes together to create our overall dataframe
IndBehav2<- rbind(FemaleOnly, MaleOnly)

#View(IndBehav2)
```

\subsubsection{Calculating the behavioral tolerance metrics needed for analyses}

We are calculating behavioral tolerance through two metrics. First is a measure of behavioral tolerance using linear mixed models using slope random effects for fishID. This allows us to extract each individuals change in activity with parasite burden and therefore allow us to calculate the tolerance (fishID slope) for each individual. This also allows us to use the random effect intercept for fishID as a measure of beahvioral vigor. The second way we could caluclate these metrics are a measure of change in behavior between two points of behavior (often referred to as point tolerance, CITATION FOR THIS).  This is calculated in teh second half of the code as the change in behavior between pre-infection and early, late, and later points of infection. 

```{r calculating the tolerance metrics, warning=FALSE}

#Calulating the behavioral tolerance metrics were interested in for our analysis

#Random slope from random effects model
#Random effects model with random slope for each fish by worm burden prior to measuring its activity. This slope is the tolerance an individual has to changing its behavior as infection increases. The slope of the behavior is the individuals behavioral vigor, or its pre-infection behavior
BehavTolLM<-lmer(Velocity~(Wormbf|fishID), IndBehav2)
summary(BehavTolLM)
#Extracting the intercept and slope from the linear mixed model
betaTol <- data.frame(coef(BehavTolLM)$fishID)
#Renaming the slope and intercept to align with our metrics
betaTol <- betaTol %>%
  rename(BehavTol = Wormbf, BehavVig = X.Intercept.)
#Creating a fishID column for combination with future dataframe
betaTol<-cbind(fishID=rownames(betaTol), betaTol)

#Calculating the mean and variance in velocity per day, per fish

#We need to split the dataframes into the differnet time periods to calculate averages since mutate isnt splitting by TrialTime. 

#creating a dataframe to calculate the mean and variance of velocity for before measurements. 

InBehavBefore<- IndBehav2%>%
  drop_na(Velocity)%>%
  filter(TrialTime == "Before")%>%
  select(-c(InfDiff, Fish, Behavdate, Duration, Distance, comp3, Totworm, rate05, rate012,rate512,rate1216,rate18end,TOD,day, linf,maxlinf, maxworm,peakday,Ddate,Rdate, CountDay))%>%
  mutate(AvgVel = mean(Velocity), VarVel = var(Velocity))%>%
  distinct(fishID, .keep_all=TRUE)
  
#creating a dataframe to calculate the mean and variance of velocity for early measurements. 

InBehavEarly<- IndBehav2%>%
  drop_na(Velocity)%>%
  filter(TrialTime == "Early")%>%
  select(-c(InfDiff, Fish, Behavdate, Duration, Distance, comp3, Totworm, rate05, rate012,rate512,rate1216,rate18end,TOD,day, linf,maxlinf, maxworm,peakday,Ddate,Rdate, CountDay))%>%
  mutate(AvgVel = mean(Velocity), VarVel = var(Velocity))%>%
  distinct(fishID, .keep_all=TRUE)


#creating a dataframe to calculate the mean and variance of velocity for Late measurements. 

InBehavLate<- IndBehav2%>%
  drop_na(Velocity)%>%
  filter(TrialTime == "Late")%>%
  select(-c(InfDiff, Fish, Behavdate, Duration, Distance, comp3, Totworm, rate05, rate012,rate512,rate1216,rate18end,TOD,day, linf,maxlinf, maxworm,peakday,Ddate,Rdate, CountDay))%>%
  mutate(AvgVel = mean(Velocity), VarVel = var(Velocity))%>%
  distinct(fishID, .keep_all=TRUE)

#creating a dataframe to calculate the mean and variance of velocity for Later measurements. 

InBehavLater<- IndBehav2%>%
  drop_na(Velocity)%>%
  filter(TrialTime == "Later")%>%
  select(-c(InfDiff, Fish, Behavdate, Duration, Distance, comp3, Totworm, rate05, rate012,rate512,rate1216,rate18end,TOD,day, linf,maxlinf, maxworm,peakday,Ddate,Rdate, CountDay))%>%
  mutate(AvgVel = mean(Velocity), VarVel = var(Velocity))%>%
  distinct(fishID, .keep_all=TRUE)

#Adding these dataframes together to create one large one

IndBehav3<-rbind(InBehavBefore, InBehavEarly, InBehavLate, InBehavLater)


#Pivoting wider so I can subtract velocity changes between periods to get change in behaivor
IndBehav4 <- IndBehav3 %>%
  group_by(TrialTime, fishID, add=TRUE)%>%
  select(-Velocity)%>%
  pivot_wider(names_from = TrialTime, values_from = c( Wormbf, Ratebf, AvgVel,VarVel))%>%
  rename(Velocity)

#subset down to infected individuals only and calculate point tolerance metrics for each timeframe. 
IndBehav4Inf <- IndBehav4 %>%
  group_by(fishID)%>%
  filter(Infection == 1) %>%
  mutate(EChBe = AvgVel_Early - AvgVel_Before,
         LChBe = AvgVel_Late - AvgVel_Before,
         LtrChBe = AvgVel_Later - AvgVel_Before,
         ERatebf = Wormbf_Early - Wormbf_Before/6,
         LRatebf = Wormbf_Late - Wormbf_Early/6,
         LtrRatebf = Wormbf_Later - Wormbf_Later/6
         )%>%
  pivot_longer(cols=starts_with("Wormbf"),
               names_to = "TrialTime",
               names_prefix = "Wormbf_",
               values_to = "Wormbf")%>%
  select(-c(AvgVel_Before, AvgVel_Early, AvgVel_Late, AvgVel_Later, Ratebf_Before,Ratebf_Early,Ratebf_Late,Ratebf_Later))%>%
  filter(fishID != 44)
#Make two dataframes with pivot longer for wormbf and velocity then merge them together on fishID and TrialTime
IndBehav4InfV<-IndBehav4 %>%
   filter(Infection == 1) %>%
  pivot_longer(cols = starts_with("AvgVel"),
    names_to = "TrialTime",
    names_prefix = "AvgVel_",
    values_to = "AvgVel"
  )%>%
  select(fishID, AvgVel, TrialTime)%>%
  filter(fishID != 44 )

#Making the growth rate columns into one column for infected individuals
IndBehavGRInf <- IndBehav4 %>%
  filter(Infection == 1) %>%
  pivot_longer(cols = starts_with("Ratebf"),
      names_to = "TrialTime",
     names_prefix = "Ratebf_",
     values_to = "Ratebf"
   )%>%
  select(fishID,TrialTime, Ratebf)

#Making the new growth rate columns into one column for infected individuals
IndBehavNGRInf <- IndBehav4Inf %>%
  select(fishID, ERatebf, LRatebf, LtrRatebf)%>%
  rename(NRatebf_Early = ERatebf, NRatebf_Late = LRatebf, NRatebf_Later = LtrRatebf)%>%
  add_column(NRatebf_Before = 0)%>%
  pivot_longer(cols = starts_with("NRatebf"),
      names_to = "TrialTime",
     names_prefix = "NRatebf_",
     values_to = "NRatebf"
   )%>%
  select(fishID,TrialTime, NRatebf)%>%
  distinct(fishID,TrialTime, .keep_all=TRUE)

#Making the Variance columns into one column for infected individuals
IndBehavVarInf <- IndBehav4Inf %>%
  select(-TrialTime)%>%
  pivot_longer(cols = starts_with("VarVel"),
      names_to = "TrialTime",
     names_prefix = "VarVel_",
     values_to = "VarVel"
   )%>%
  select(fishID,TrialTime, VarVel)%>%
  distinct(fishID,TrialTime, .keep_all=TRUE)


#Merging the tolerance and vigor metrics 
IndBehav4Inf1<-merge(IndBehav4Inf, betaTol, by.x="fishID", by.y = "fishID")

#Merge the two dataframes together based on the fishID and TrialTime to get all variables we want into one dataframe
IndBehav5Inf<-merge(IndBehav4InfV, IndBehav4Inf1, by.x=c("fishID", "TrialTime"), by.y=c("fishID", "TrialTime"), all=TRUE)
IndBehav5Inf2<-merge(IndBehav5Inf, IndBehavGRInf, by.x=c("fishID", "TrialTime"), by.y=c("fishID", "TrialTime"), all=TRUE)
IndBehav5Inf3<-merge(IndBehav5Inf2, IndBehavNGRInf, by.x=c("fishID", "TrialTime"), by.y=c("fishID", "TrialTime"), all=TRUE)
IndBehav5Inf4<-merge(IndBehav5Inf3, IndBehavVarInf, by.x=c("fishID", "TrialTime"), by.y=c("fishID", "TrialTime"), all=TRUE)
#Subset down to uninfected individuals and calculate their point tolerance metrics
#Creating a dataframe for calculating the tolerance metrics and then pivot longer by Wormbf
IndBehav3Unf <- IndBehav3 %>%
  filter(Infection == 0)%>%
  mutate(EChBe = Velocity_Early - Velocity_Before,
         LChBe = Velocity_Late - Velocity_Before,
         LtrChBe = Velocity_Later - Velocity_Before,
         ERatebf = Wormbf_Early - Wormbf_Before/5,
         LRatebf = Wormbf_Late - Wormbf_Early/5,
         LtrRatebf = Wormbf_Later - Wormbf_Later/5,
         )%>%
  group_by(fishID)%>%
  pivot_longer(cols = starts_with(c("Wormbf")),
      names_to = "TrialTime",
     names_prefix = c("Wormbf_"),
     values_to="Wormbf"
   )%>%
  select(-c(Velocity_Before, Velocity_Early, Velocity_Late, Velocity_Later, Ratebf_Before,Ratebf_Early,Ratebf_Late,Ratebf_Later))


#Creating a second dataframe where we pivot longer for velocity
IndBehav3UnfV <- IndBehav3 %>%
  filter(Infection == 0)%>%
  pivot_longer(cols = starts_with("Velocity"),
  names_to = "TrialTime",
  names_prefix = "Velocity_",
  values_to = "Velocity"
  )%>%
  select(fishID, TrialTime, Velocity)


#Making the growth rate columns into one column for uninfected individuals
IndBehavGRUnf <- IndBehav3 %>%
  filter(Infection == 0) %>%
  pivot_longer(cols = starts_with("Ratebf"),
      names_to = "TrialTime",
     names_prefix = "Ratebf_",
     values_to = "Ratebf"
   )%>%
  select(fishID,TrialTime, Ratebf)

#Merge the two different datasets together
IndBehav4Unf<-merge(IndBehav3UnfV, IndBehav3Unf, by.x=c("fishID", "TrialTime"), by.y=c("fishID", "TrialTime"), all=TRUE)
IndBehav4Unf2<-merge(IndBehav4Unf, IndBehavGRUnf, by.x=c("fishID", "TrialTime"), by.y=c("fishID", "TrialTime"), all=TRUE)

#Creating the BehavTol and BehavVig columns for the uninfected and calculating the metric for uninfected individual so we can combine these dataframes together
IndBehav4Unf2 <- IndBehav4Unf2 %>%
  group_by(fishID)%>%
  add_column(BehavTol=0, BehavVig=0)


#Rbinding our infected and uninfected dataframes together into one overall dataframe
IndBehav4<-rbind(IndBehav4Unf2,IndBehav4Inf2)

#Remove some unwanted columns to clean up dataset in preparation for analysis
IndBehav4 <- IndBehav4%>%
  select(-c(day, TOD,linf,maxlinf, maxworm,peakday,Ddate,Rdate, CountDay))

View(IndBehav4)

```

Now we have to decided how best to calculate whether an individual has recovered. In the experiment only had 13 fish recover from infection totally, therefore we are thinking of another metric that indicates a fish is in recovery during infection. 

\subsection{Description of data, structure, and type}

\textbf{fishID}: The individual ID of each fish used in the trial. \newline
\textbf{TrialTime}: The point of infection where each behavior trial happened. Early - early infection (6 days), Late - late in infection (12 days), Later - later in infection (18 days)\newline
\textbf{Velocity}: The average velocity of each fish during their behavior trial ($cm/s$) \newline
\textbf{Sex}: The sex of the individual. F - female, M - male \newline
\textbf{PreWeight}: The weight of the individual prior to their fist behavior trial and pre-infection. ($grams$)  \newline
\textbf{PreLength}: The length of the individual prior to their first behavior trial and pre-infection ($mm$) \newline
\textbf{Treatment}: What treatment the fish received prior to their first behavior and pre-infection. VIE - visible implant elastomer implant, UNTOUCHED - control individual, received no injection nor implant ($mm$) \newline
\textbf{Infection}: Whether or not the individual was infected with Gyrodactylus turnbulli. 1 - infected, 0 - uninfected \newline
\textbf{LateWeight}: The weight of the individual after their final behavior trial and after infection. ($grams$) \newline
\textbf{wormJump}: The number of worms that jumped from the donor fish to the trial fish during manual infections. \newline
\textbf{AUC}: The area under the curve of infection over the total infection trajectory for each individual. \newline
\textbf{Recov}: Whether or not the individual recovered from infection during the experimental trial. 1 - recovered, 0 - did not recover \newline
\textbf{Died}: Whether or not the individual Died from infection during the experimental trial. 1 - Died, 0 - did not died \newline
\textbf{PreSMI}: The body condition of the individual prior to their first behavioral trial and pre infection. ($mm/g$) \newline
\textbf{LateSMI}: The body condition of the individual after to their last behavioral trial and after infection. ($mm/g$) \newline
\textbf{EChBe}: The change in velocity between the early infection behavioral trial and the before trial. ($cm/s$) \newline
\textbf{LChBe}: The change in velocity between the Late infection behavioral trial and the before trial. ($cm/s$) \newline
\textbf{LtrChBe}: The change in velocity between the Later infection behavioral trial and the before trial. ($cm/s$) \newline
\textbf{Wormbf}:The number of worms on the fish prior to each trial. \newline
\textbf{Ratebf}:The rate of growth of worms on the fish prior to each trial. \newline
\textbf{BehavTol}:The behavioral tolerance of each individual calculated as the slope of a random effect model where fishID is the random effect term.  \newline
\textbf{BehavVig}:The behavioral Vigor of each individual calculated as the intercept of a random effect model where fishID is the random effect term \newline